---
title: "Sonoma County basins, sanity check"
output: 
  html_document:
    code_folding: hide
---

Last generated: `r Sys.time()`. <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%")

library(tidyverse)
library(sf)
library(here)
library(fs)
library(mapview)

# GSA spatial data
b118_path <- path(data_path, "general/b118/i08_B118_v6-1.shp") 

# gsa spatial data: petaluma, sonoma valley, santa rosa plain
son <- f_load_b118_basin(b118_path, "NAPA-SONOMA VALLEY - SONOMA VALLEY")
pet <- f_load_b118_basin(b118_path, "PETALUMA VALLEY")
srp <- f_load_b118_basin(b118_path, "SANTA ROSA VALLEY - SANTA ROSA PLAIN")

# borders for mapview
son_border <- st_cast(son, "MULTILINESTRING")
pet_border <- st_cast(pet, "MULTILINESTRING")
srp_border <- st_cast(srp, "MULTILINESTRING")

# groundwater budgets from groundwater flow models
# pumping per user based on models, postprocessed by 01_avg_annual...R, 
pump <- path(data_path, "tables/pump.csv") %>% 
  read_csv()

# SON parcel database
pson <- read_rds(path(data_path, "data_output/son_parcel_complete.rds")) %>% 
  mutate(
    type = ifelse(
      UseCode_Category != "Agricultural" | is.na(UseCode_Category), 
      "M&I plus domestic", "agriculture"),
    basin = "Sonoma Valley")

# PET parcel database
ppet <- read_rds(path(data_path, "data_output/pet_parcel_complete.rds")) %>% 
  mutate(
    type = case_when(
      UseCode_Category == "Agricultural" ~ "agriculture",
      UseCode_Category == "Commercial" ~ "M&I", 
      UseCode_Category %in% c("MultiFamily", "Residential") ~ "domestic", 
      TRUE ~ UseCode_Category),
    basin = "Petaluma Valley")

# SRP parcel database
srp_use_ag  <- c("Dry Farm", "Irrigated Farm")
srp_use_mid <- c("Recreational","Institutional","Miscellaneous","Industrial",
                 "Government","Commercial","Residential")
psrp <- read_rds(path(data_path, "data_output/srp_parcel_complete.rds")) %>% 
  mutate(
    # clean up and sort out use codes
    UseCode_Category = str_remove_all(UseCode_Category, " Properties"),
    type = case_when(
      UseCode_Category %in% srp_use_ag ~ "agriculture",
      UseCode_Category %in% srp_use_mid ~ "M&I plus domestic", 
      TRUE ~ UseCode_Category),
    basin = "Santa Rosa Plain")
```

# Rebucketizing

`UseCodeCategory` "buckets", although simple, inadvertently combine water uses. For example, a parcel with "Commercial" `UseCodeCategory` may contain agricultural acerage, and have predominately agricultural water use. Here, we experiment with different ways to aggregate the data (e.g., "rebucketize"). First, using `UseCodeCategory`:

```{r}
srp_usecode_bucket <- psrp %>% 
  st_drop_geometry() %>% 
  group_by(type) %>% 
  summarise(pumpage_af = sum(Total_Groundwater_Use_Ac_Ft, 
                             na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(desc(pumpage_af)) %>% 
  filter(!is.na(type))

son_usecode_bucket <- pson %>% 
  st_drop_geometry() %>% 
  group_by(type) %>% 
  summarise(pumpage_af = sum(Total_Groundwater_Use_Ac_Ft, 
                             na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(desc(pumpage_af))

pet_usecode_bucket <- ppet %>% 
  st_drop_geometry() %>% 
  group_by(type) %>% 
  summarise(pumpage_af = sum(Total_Groundwater_Use_Ac_Ft, 
                             na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(desc(pumpage_af)) %>% 
  filter(!is.na(type))

# results from bucket 1
srp_usecode_bucket
son_usecode_bucket
pet_usecode_bucket
```

Next, buckets based on raw water use:

```{r}
srp_raw_bucket <- psrp %>% 
  st_drop_geometry() %>% 
  summarise(
    agriculture = sum(Ag_GW_Use_Ac_Ft),
    residential = sum(Res_GW_Use_Ac_Ft),
    commercial  = sum(Commercial_GW_Use_Ac_Ft),
    school_golf = sum(School_Golf_GW_Use_Ac_Ft),
    urban_irrigation = sum(Urban_Irrigation_GW_Use_Ac_Ft)
  ) %>% 
  mutate(
    agriculture = agriculture + school_golf,
    `M&I plus domestic` = residential + commercial + urban_irrigation,
    total = agriculture + `M&I plus domestic`
  )

son_raw_bucket <- pson %>% 
  st_drop_geometry() %>% 
  summarise(
    agriculture = sum(Ag_GW_Use_Ac_Ft),
    residential = sum(Res_GW_Use_Ac_Ft),
    commercial  = sum(Commercial_GW_Use_Ac_Ft),
    school_golf = sum(School_Golf_GW_Use_Ac_Ft),
    urban_irrigation = sum(Urban_Irrigation_GW_Use_Ac_Ft)
  ) %>% 
  mutate(
    agriculture = agriculture + school_golf,
    `M&I plus domestic` = residential + commercial + urban_irrigation,
    total = agriculture + `M&I plus domestic`
  )

pet_raw_bucket <- ppet %>% 
  st_drop_geometry() %>% 
  summarise(
    agriculture = sum(Ag_GW_Use_Ac_Ft),
    residential = sum(Res_GW_Use_Ac_Ft),
    commercial  = sum(Commercial_GW_Use_Ac_Ft),
    school_golf = sum(School_Golf_GW_Use_Ac_Ft),
    urban_irrigation = sum(Urban_Irrigation_GW_Use_Ac_Ft)
  ) %>% 
  mutate(
    agriculture = agriculture + school_golf,
    `M&I`    = commercial + urban_irrigation,
    domestic = residential, 
    total    = agriculture + `M&I` + domestic
  )

# results from bucket 2
srp_raw_bucket %>% 
  select(agriculture, `M&I plus domestic`) %>%
  pivot_longer(everything(), names_to = "type", values_to = "pumpage_af")

son_raw_bucket %>% 
  select(agriculture, `M&I plus domestic`) %>%
  pivot_longer(everything(), names_to = "type", values_to = "pumpage_af")

pet_raw_bucket %>% 
  select(agriculture, `M&I`, domestic) %>%
  pivot_longer(everything(), names_to = "type", values_to = "pumpage_af")
```


Buckets based on proportion of ag land, where >= 50% ag is reclassified as "ag" were tried but don't work well.

```{r, eval = FALSE}
pson %>% 
  st_drop_geometry() %>% 
  mutate(is_ag = ifelse(
    (Total_Crop_Area_Ac / LandSizeAcres) >= 0.5, TRUE, FALSE)) %>% 
  group_by(is_ag) %>% 
  summarise(gw_use = sum(Total_Groundwater_Use_Ac_Ft))
  
ppet %>% 
  st_drop_geometry() %>% 
  mutate(is_ag = ifelse(
    (Total_Crop_Area_Ac / LandSizeAcres) >= 0.50, TRUE, FALSE)) %>% 
  group_by(is_ag) %>% 
  summarise(gw_use = sum(Total_Groundwater_Use_Ac_Ft))
```

Finally, buckets based on raw water use where non-zero commercial and residential water uses on "Ag" Accessor use land are reclassified as Ag water uses. 

```{r}
srp_raw_bucket2 <- psrp %>% 
  st_drop_geometry() %>% 
  mutate(
    Ag_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      Ag_GW_Use_Ac_Ft + Res_GW_Use_Ac_Ft + Commercial_GW_Use_Ac_Ft, 
      Ag_GW_Use_Ac_Ft),
    Res_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      0, 
      Res_GW_Use_Ac_Ft),
    Commercial_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      0, 
      Commercial_GW_Use_Ac_Ft)
  ) %>% 
  summarise(
    agriculture = sum(Ag_GW_Use_Ac_Ft),
    residential = sum(Res_GW_Use_Ac_Ft),
    commercial  = sum(Commercial_GW_Use_Ac_Ft),
    school_golf = sum(School_Golf_GW_Use_Ac_Ft),
    urban_irrigation = sum(Urban_Irrigation_GW_Use_Ac_Ft)
  ) %>% 
  mutate(
    agriculture = agriculture + school_golf,
    `M&I plus domestic` = residential + commercial + urban_irrigation,
    total = agriculture + `M&I plus domestic`
  )

son_raw_bucket2 <- pson %>% 
  st_drop_geometry() %>% 
  mutate(
    Ag_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      Ag_GW_Use_Ac_Ft + Res_GW_Use_Ac_Ft + Commercial_GW_Use_Ac_Ft, 
      Ag_GW_Use_Ac_Ft),
    Res_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      0, 
      Res_GW_Use_Ac_Ft),
    Commercial_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      0, 
      Commercial_GW_Use_Ac_Ft)
  ) %>% 
  summarise(
    agriculture = sum(Ag_GW_Use_Ac_Ft),
    residential = sum(Res_GW_Use_Ac_Ft),
    commercial  = sum(Commercial_GW_Use_Ac_Ft),
    school_golf = sum(School_Golf_GW_Use_Ac_Ft),
    urban_irrigation = sum(Urban_Irrigation_GW_Use_Ac_Ft)
  ) %>% 
  mutate(
    agriculture = agriculture + school_golf,
    `M&I plus domestic` = residential + commercial + urban_irrigation,
    total = agriculture + `M&I plus domestic`
  )

pet_raw_bucket2 <- ppet %>% 
  st_drop_geometry() %>% 
  mutate(
    Ag_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      Ag_GW_Use_Ac_Ft + Res_GW_Use_Ac_Ft + Commercial_GW_Use_Ac_Ft, 
      Ag_GW_Use_Ac_Ft),
    Res_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      0, 
      Res_GW_Use_Ac_Ft),
    Commercial_GW_Use_Ac_Ft = ifelse(
      type == "agriculture", 
      0, 
      Commercial_GW_Use_Ac_Ft)
  ) %>% 
  summarise(
    agriculture = sum(Ag_GW_Use_Ac_Ft, na.rm = TRUE),
    residential = sum(Res_GW_Use_Ac_Ft, na.rm = TRUE),
    commercial  = sum(Commercial_GW_Use_Ac_Ft, na.rm = TRUE),
    school_golf = sum(School_Golf_GW_Use_Ac_Ft, na.rm = TRUE),
    urban_irrigation = sum(Urban_Irrigation_GW_Use_Ac_Ft, na.rm = TRUE)
  ) %>% 
  mutate(
    agriculture = agriculture + school_golf,
    `M&I`    = commercial + urban_irrigation,
    domestic = residential, 
    total    = agriculture + `M&I` + domestic
  )

# results from bucket 2
srp_raw_bucket2 %>% 
  select(agriculture, `M&I plus domestic`) %>%
  pivot_longer(everything(), names_to = "type", values_to = "pumpage_af")

son_raw_bucket2 %>% 
  select(agriculture, `M&I plus domestic`) %>%
  pivot_longer(everything(), names_to = "type", values_to = "pumpage_af")

pet_raw_bucket2 %>% 
  select(agriculture, `M&I`, domestic) %>%
  pivot_longer(everything(), names_to = "type", values_to = "pumpage_af")
```


# Crop acres per basin and water use

Here, we summarize crop acres per basin and consumptive demand to compare with groundwater flow models. Importantly, because a parcel may have multiple crops and multiple sources of surface and recycled water in addition to groundwater, reported demands are not adjusted for the actual groundwater consumed per crop. 

This calculation is most important for PET [& SRP], where parcel-estimated agricultural water use is below average modeled agricultural groundwater use by around 600-700 AF/yr. 

### SRP

```{r}
psrp_crop_acres <- psrp %>% 
  st_drop_geometry() %>% 
  summarise(
    across(contains("Area_Ac"), ~sum(.x))
  ) %>% 
  pivot_longer(everything(), names_to = "crop", values_to = "acres") %>% 
  mutate(crop = str_remove_all(crop, "_Area_Ac"))

psrp_crop_demand <- psrp %>% 
  st_drop_geometry() %>% 
  summarise(
    across(contains("_Rate"), ~sum(.x))
  ) %>% 
  pivot_longer(everything(), names_to = "name", values_to = "AF_yr")

# reshuffling needed for SRP
psrp_crop_demand <- bind_rows(slice(psrp_crop_demand, 2:10),
                              slice(psrp_crop_demand, 1))

bind_cols(psrp_crop_acres, psrp_crop_demand) %>% 
  select(-name)
```

### SON

```{r}
pson_crop_acres <- pson %>% 
  st_drop_geometry() %>% 
  summarise(
    across(contains("Area_Ac"), ~sum(.x))
  ) %>% 
  pivot_longer(everything(), names_to = "crop", values_to = "acres") %>% 
  mutate(crop = str_remove_all(crop, "_Area_Ac"))

pson_crop_demand <- pson %>% 
  st_drop_geometry() %>% 
  summarise(
    across(contains("_Rate"), ~sum(.x))
  ) %>% 
  pivot_longer(everything(), names_to = "name", values_to = "AF_yr")

bind_cols(pson_crop_acres, pson_crop_demand) %>% 
  select(-name)
```

### PET

```{r}
ppet_crop_acres <- ppet %>% 
  st_drop_geometry() %>% 
  summarise(
    across(contains("Area_Ac"), ~sum(.x))
  ) %>% 
  pivot_longer(everything(), names_to = "crop", values_to = "acres") %>% 
  mutate(crop = str_remove_all(crop, "_Area_Ac"))

ppet_crop_demand <- ppet %>% 
  st_drop_geometry() %>% 
  summarise(
    across(contains("_Rate"), ~sum(.x))
  ) %>% 
  pivot_longer(everything(), names_to = "name", values_to = "AF_yr")

bind_cols(ppet_crop_acres, ppet_crop_demand) %>% 
  select(-name)
```


# Rural residential parcels

We provide count and location of rural residential parcels per basin, based on non-zero Residential GW Use.

First focus on PET because the parcel bucket model's domestic use seems too high. `r nrow(filter(ppet, Res_GW_Use_Ac_Ft > 0))` parcels with non-zero Residential water use (from Accessor codes) contribute to the `r sum(ppet$Res_GW_Use_Ac_Ft)` of domestic groundwater use. 

```{r}
# pet_rr <- ppet %>% 
#   filter(UseCode_Category == "Residential" & 
#            Public_Water_Connection == "No") 
# 
# sum(pet_rr$Total_Groundwater_Use_Ac_Ft)

ppet %>% 
  filter(Res_GW_Use_Ac_Ft > 0) %>% 
  mapview(zcol = "Res_GW_Use_Ac_Ft") + 
  mapview(pet_border)

sum(ppet$Res_GW_Use_Ac_Ft) # acre-feet
```


# Public water supply systems and water uses

Water use is calculated at parcels with a defined public water connection, and are grouped by drinking water service area. Groundwater use is not included, thus uses reported only represent the water delivered by the system.


### SRP

```{r}
# what water systems *should* we have water use data for?
psrp %>% 
  st_drop_geometry() %>% 
  count(CA_DrinkingWater_SvcArea_Name, sort = TRUE)

# what water systems do we *actually* have water use data for?
wsu_srp <- psrp %>% 
  st_drop_geometry() %>% 
  filter(Public_Water_Connection == "Yes", 
         CA_DrinkingWater_SvcArea_Name != "Not In Public Water System") %>% 
  group_by(CA_DrinkingWater_SvcArea_Name) %>% 
  summarise(use_af_parcel = 
              sum(Res_W_Use_Assessor_Ac_Ft, 
                  Commercial_W_Use_Assessor_Ac_Ft, na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(desc(use_af_parcel)) 
wsu_srp
```


## SON

```{r}
# what water systems *should* we have water use data for?
pson %>% 
  st_drop_geometry() %>% 
  count(CA_DrinkingWater_SvcArea_Name, sort = TRUE)

# what water systems do we *actually* have water use data for?
wsu_son <- pson %>% 
  st_drop_geometry() %>% 
  filter(Public_Water_Connection == "Yes",
         !is.na(CA_DrinkingWater_SvcArea_Name)) %>%
  group_by(CA_DrinkingWater_SvcArea_Name) %>% 
  summarise(use_af_parcel = 
              sum(Res_W_Use_Assessor_Ac_Ft, 
                  Commercial_W_Use_Assessor_Ac_Ft, na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(desc(use_af_parcel))
wsu_son
```


## PET

```{r}
# what water systems *should* we have water use data for?
ppet %>% 
  st_drop_geometry() %>% 
  count(CA_DrinkingWater_SvcArea_Name, sort = TRUE)

# what water systems do we *actually* have water use data for?
wsu_pet <- ppet %>% 
  st_drop_geometry() %>% 
  filter(Public_Water_Connection == "Yes",
         !is.na(CA_DrinkingWater_SvcArea_Name)) %>%
  group_by(CA_DrinkingWater_SvcArea_Name) %>% 
  summarise(use_af_parcel = 
              sum(Res_W_Use_Assessor_Ac_Ft, 
                  Commercial_W_Use_Assessor_Ac_Ft, na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(desc(use_af_parcel))
wsu_pet
```

Combine model estimated water data (from Accessor rates and water system boundaries as calculated above) with data reported to the state water Board.

```{r}
# annual water use from the Water Board
wu <- xlsx::read.xlsx(
  path(data_path, "general/pws_water_use",
       "2020 EAR Water Suppy and Deilvery data for Sonoma District PWS.xlsx"), 
  sheetIndex = 1, colIndex = c(1, 7, 99)) %>% 
  rename(pwsid = PwsID, use = WPAnnualTotal, unit = WPUnitsofMeasure) %>% 
  mutate(
    # convert all units to AF
    use_af_board = case_when(
      unit == "MG"  ~ use * 3.68558063,
      unit == "G"   ~ use * 3.06889e-6,
      unit == "CCF" ~ use * 0.00229569,
      unit == "AF"  ~ use * 1),
    unit = "AF"
  ) %>% 
  select(-use)

# water service areas in California
rm_pwsid <- c("CA2110003")
wsa <- path(data_path, "general", "water_system_boundaries",
            "SABL_Public_083121/SABL_Public_083121.shp") %>% 
  st_read() %>% 
  st_transform(epsg) %>% 
  st_intersection(st_union(son, pet, srp)) %>% 
  select(CA_DrinkingWater_SvcArea_Name = WATER_SY_1,
         pwsid = SABL_PWSID) %>% 
  left_join(wu) %>%
  filter(!is.na(use_af_board) & ! pwsid %in% rm_pwsid)

# attach water use
wu <- arrange(wsa, desc(use_af_board)) %>% 
  left_join(bind_rows(wsu_srp, wsu_son, wsu_pet)) %>%
  mutate(name = ifelse(
    use_af_parcel > 500,
    str_wrap(CA_DrinkingWater_SvcArea_Name, 13),
    NA)
  )
  
p1 <- wu %>% 
  ggplot(aes(use_af_parcel, use_af_board, 
             text = CA_DrinkingWater_SvcArea_Name)) + 
  geom_point(alpha = 0.5) + 
  geom_abline(slope = 1, linetype = "dashed") +
  ggrepel::geom_label_repel(aes(label = name), size = 2,
                            fill = "white", alpha = 1) + 
  annotate("rect", xmin = -10, xmax = 100, ymin = -10, ymax = 100,
           alpha = .1, color = "red") +
  coord_cartesian(ylim = c(0, 4000), xlim = c(0, 4000)) +
  scale_y_continuous(expand = c(0.01, 0.01)) +
  scale_x_continuous(expand = c(0.01, 0.01)) +
  labs(x = "Water use (Parcel)", y = "Water use (Board)",
       title = "Comparison of Annual Water Use (AF)",
       subtitle = "Parcel (estimated) and Water Board (reported)") +
  cowplot::theme_cowplot()

p2 <- wu %>% 
  ggplot(aes(use_af_parcel, use_af_board)) + 
  geom_point(alpha = 0.5) + 
  geom_abline(slope = 1, linetype = "dashed") +
  labs(x = "", y = "") +
  coord_cartesian(ylim = c(0, 100), xlim = c(0, 100)) +
  cowplot::theme_cowplot()  + 
  theme(plot.background = element_rect(color = "red"), 
        plot.margin = unit(c(10,10,-10,-10), units = "points"))

p1 + inset_element(p2, 0.05, 0.6, 0.45, 1)
plotly::ggplotly(p1, tooltip = c("text", "x", "y"))
```

